19Q2

### 2019-07-02 19Q2.Rmd
update prot Rds and ht w/ better dir and NA annots

gdsc_signif_prot_res <- readRDS("./data_munging/rds/gdsc_signif_protein_changes_resistance.rds")
gdsc_signif_prot_res$Direction_Test <- "Resistance"
gdsc_signif_prot_sen <- readRDS("./data_munging/rds/gdsc_signif_protein_changes_sensitivity.rds")
gdsc_signif_prot_sen$Direction_Test <- "Sensitivity"
gdsc_signif_prot_dir <- bind_rows(gdsc_signif_prot_res, gdsc_signif_prot_sen)
```

@@

ccle_signif_prot_4plot <- merge(merge(ccle_signif_prot, unique(select(ccle_prot, Drug, Gene_Protein_Change, Direction_Curated)), by = c("Drug", "Gene_Protein_Change"), all.x = TRUE), expand.grid("Drug" = unique(ccle_signif_prot$Drug), "Gene_Protein_Change" = gene_prots), by = c("Drug", "Gene_Protein_Change"), all = TRUE)
ctrp_signif_prot_4plot <- merge(merge(ctrp_signif_prot, unique(select(ctrp_prot, Drug, Gene_Protein_Change, Direction_Curated)), by = c("Drug", "Gene_Protein_Change"), all.x = TRUE), expand.grid("Drug" = unique(ctrp_signif_prot$Drug), "Gene_Protein_Change" = gene_prots), by = c("Drug", "Gene_Protein_Change"), all = TRUE)
gdsc_signif_prot_4plot <- merge(merge(gdsc_signif_prot, unique(select(gdsc_prot, Drug, Gene_Protein_Change, Direction_Curated)), by = c("Drug", "Gene_Protein_Change"), all.x = TRUE), expand.grid("Drug" = unique(gdsc_signif_prot$Drug), "Gene_Protein_Change" = gene_prots), by = c("Drug", "Gene_Protein_Change"), all = TRUE)

@@

```{r wilcoxon_heatmap_function}
MakeWilcoxonHeatmap <- function(signif_df, feature, row_names, col_title, color, legend_title, legend_breaks, row_title_txt, radius) {
  if(!is.na(str_match(string = tolower(feature), pattern = "mut"))) {
    signif_grid <- dcast(signif_df, Gene ~ Drug, value.var = "p_log10", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
    signif_grid <- signif_grid[g2p_gene_order, intersect(g2p_drug_order, colnames(signif_grid))]
    signif_grid <- signif_grid[complete.cases(signif_grid), ]
    signif_grid[signif_grid < -log10(0.05)] <- NA

    signif_grid_annot <- dcast(signif_df, Gene ~ Drug, value.var = "InG2P", fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
    signif_grid_annot <- signif_grid_annot[g2p_gene_order, intersect(g2p_drug_order, colnames(signif_grid))]
    signif_grid_annot <- signif_grid_annot[complete.cases(signif_grid_annot), ]
    signif_grid_annot[signif_grid_annot == "No"] <- ""
    signif_grid_annot[signif_grid_annot == "Yes"] <- "O"
  }
  if(!is.na(str_match(string = tolower(feature), pattern = "prot"))) {
    signif_grid <- dcast(signif_df, Gene_Protein_Change ~ Drug, value.var = "p_log10", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene_Protein_Change)
    signif_grid <- signif_grid[, intersect(g2p_drug_order, colnames(signif_grid))]
    # signif_grid <- signif_grid[complete.cases(signif_grid), ]
    signif_grid[signif_grid < -log10(0.05)] <- NA

    signif_grid_annot <- dcast(signif_df, Gene_Protein_Change ~ Drug, value.var = "Direction_Curated", fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene_Protein_Change)
    signif_grid_annot <- signif_grid_annot[, intersect(g2p_drug_order, colnames(signif_grid))]
    # signif_grid_annot <- signif_grid_annot[complete.cases(signif_grid_annot), ]
    signif_grid_annot[signif_grid_annot == "resistance"] <- "X"
    signif_grid_annot[signif_grid_annot == "sensitivity"] <- "O"
  }
  if(!is.na(str_match(string = tolower(feature), pattern = "fus"))) {
    signif_grid <- dcast(signif_df, FusionPair ~ Drug, value.var = "p_log10", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-FusionPair)
    signif_grid <- signif_grid[, intersect(g2p_drug_order, colnames(signif_grid))]
    signif_grid <- signif_grid[complete.cases(signif_grid), ]
    signif_grid[signif_grid < -log10(0.05)] <- NA

    signif_grid_annot <- dcast(signif_df, FusionPair ~ Drug, value.var = "InG2P", fill = 0) %>% `rownames<-`(.[,1]) %>% select(-FusionPair)
    signif_grid_annot <- signif_grid_annot[, intersect(g2p_drug_order, colnames(signif_grid))]
    signif_grid_annot <- signif_grid_annot[complete.cases(signif_grid_annot), ]
    signif_grid_annot[signif_grid_annot == "No"] <- ""
    signif_grid_annot[signif_grid_annot == "Yes"] <- "O"
  }
  col_fun <- colorRamp2(c(legend_breaks[1], legend_breaks[length(legend_breaks)]), c("gray97", color))
  # Legend properties list
  ht_legend_list <- list(title = legend_title, title_position = "topleft", legend_direction = "horizontal", legend_width = unit(2, "in"), border = "gray70", at = legend_breaks, labels = legend_breaks)
  ht <- Heatmap(signif_grid,
                col = col_fun,
                # na_col = "gray97",
                na_col = "white",
                rect_gp = gpar(col = "gray70"),
                heatmap_legend_param = ht_legend_list,
                cluster_rows = FALSE, 
                cluster_columns = FALSE, 
                show_row_names = as.logical(row_names),
                row_names_side = "left",
                row_names_gp = gpar(fontsize = 10),
                column_names_side = "bottom",
                column_names_gp = gpar(fontsize = 10),
                show_column_names = FALSE,
                bottom_annotation = HeatmapAnnotation(
                  text = anno_text(colnames(signif_grid),
                                   rot = 45,
                                   gp = gpar(fontsize = 11),
                                   offset = unit(1, "npc"),
                                   just = "right"),
                  annotation_height = unit(0.85, "in")),
                  # annotation_height = max_text_width(colnames(signif_grid))),
                row_title = row_title_txt,
                row_title_gp = gpar(fontface = "bold", fontsize = 14),
                column_title = col_title, 
                column_title_gp = gpar(fontface = "bold", fontsize = 14),
                column_title_side = "bottom",
                cell_fun = function(j, i, x, y, width, height, fill) {
                  # grid.text(signif_grid_annot[i, j], x, y, gp = gpar(fontsize = 20))
                  if(signif_grid_annot[i, j] == "O") {
                    grid.circle(x = x - width*0.20, y = y - height*0.20, r = radius, default.units = "npc", gp = gpar(fill = "white", col = "black"))
                  }
                  if(signif_grid_annot[i, j] == "X") {
                    grid.circle(x = x - width*0.20, y = y - height*0.20, r = radius, default.units = "npc", gp = gpar(fill = "black", col = "black"))
                  }
                }
                )
  return(ht)
}
```

@@

ht_ccle2 <- MakeWilcoxonHeatmap(signif_df = ccle_signif_prot_4plot, feature = "protein changes", legend_title = "CCLE -log10(p-value)                                                                     ", legend_breaks = seq(0, 15, by = 3), row_names = TRUE, color = "darkgreen", col_title = "", row_title_txt = "Gene:Protein Change", radius = 0.03)
ht_ctrp2 <- MakeWilcoxonHeatmap(signif_df = ctrp_signif_prot_4plot, feature = "protein changes", legend_title = "CTRP -log10(p-value)                                                                    ", legend_breaks = seq(0, 15, by = 3), row_names = FALSE, color = "steelblue4", col_title = "Drugs", row_title_txt = "Gene:Protein Change", radius = 0.015)
ht_gdsc2 <- MakeWilcoxonHeatmap(signif_df = gdsc_signif_prot_4plot, feature = "protein changes", legend_title = "GDSC -log10(p-value)", legend_breaks = seq(0, 8, by = 2), row_names = FALSE, color = "turquoise4", col_title = "", row_title_txt = "Gene:Protein Change", radius = 0.015)
png(filename = "./plots/manuscript/drug_prot_heatmap.png", width = 13, height = 5, units = "in", res = 500)
draw(ht_ccle2 + ht_ctrp2 + ht_gdsc2, gap = unit(0.25, "in"), heatmap_legend_side = "top")
dev.off()
```

>>>>

gdsc_signif_prot_res <- readRDS("./data_munging/rds/gdsc_signif_protein_changes_resistance.rds")
gdsc_signif_prot_res$Direction_Test <- "Resistance"
gdsc_signif_prot_sen <- readRDS("./data_munging/rds/gdsc_signif_protein_changes_sensitivity.rds")
gdsc_signif_prot_sen$Direction_Test <- "Sensitivity"
gdsc_signif_prot_dir <- bind_rows(gdsc_signif_prot_res, gdsc_signif_prot_sen)

ccle_prot_res <- readRDS("./data_munging/rds/ccle_protein_changes_resistance.rds")
ccle_prot_sen <- readRDS("./data_munging/rds/ccle_protein_changes_sensitivity.rds")
ctrp_prot_res <- readRDS("./data_munging/rds/ctrp_protein_changes_resistance.rds")
ctrp_prot_sen <- readRDS("./data_munging/rds/ctrp_protein_changes_sensitivity.rds")
gdsc_prot_res <- readRDS("./data_munging/rds/gdsc_protein_changes_resistance.rds")
gdsc_prot_sen <- readRDS("./data_munging/rds/gdsc_protein_changes_sensitivity.rds")
```

@@

# all_directions <- unique(rbind(select(ccle_prot, Gene_Protein_Change, Drug, Direction_Curated), select(ctrp_prot, Gene_Protein_Change, Drug, Direction_Curated), select(gdsc_prot, Gene_Protein_Change, Drug, Direction_Curated)))
# saveRDS(all_directions, "./data_munging/rds/protein_changes_directions.rds")
all_directions <- readRDS("./data_munging/rds/protein_changes_directions.rds")

ccle_signif_prot_4plot <- merge(merge(merge(ccle_signif_prot, unique(select(ccle_prot, Drug, Gene_Protein_Change)), by = c("Drug", "Gene_Protein_Change"), all.x = TRUE), expand.grid("Drug" = unique(ccle_signif_prot$Drug), "Gene_Protein_Change" = gene_prots), by = c("Drug", "Gene_Protein_Change"), all = TRUE), all_directions,  by = c("Drug", "Gene_Protein_Change"), all.x = TRUE)
ctrp_signif_prot_4plot <- merge(merge(merge(ctrp_signif_prot, unique(select(ctrp_prot, Drug, Gene_Protein_Change)), by = c("Drug", "Gene_Protein_Change"), all.x = TRUE), expand.grid("Drug" = unique(ctrp_signif_prot$Drug), "Gene_Protein_Change" = gene_prots), by = c("Drug", "Gene_Protein_Change"), all = TRUE), all_directions,  by = c("Drug", "Gene_Protein_Change"), all.x = TRUE)
gdsc_signif_prot_4plot <- merge(merge(merge(gdsc_signif_prot, unique(select(gdsc_prot, Drug, Gene_Protein_Change)), by = c("Drug", "Gene_Protein_Change"), all.x = TRUE), expand.grid("Drug" = unique(gdsc_signif_prot$Drug), "Gene_Protein_Change" = gene_prots), by = c("Drug", "Gene_Protein_Change"), all = TRUE), all_directions,  by = c("Drug", "Gene_Protein_Change"), all.x = TRUE)

@@

```{r wilcoxon_heatmap_function}
MakeWilcoxonHeatmap <- function(signif_df, feature, row_names, col_title, color, legend_title, legend_breaks, row_title_txt, radius, na_color) {
  if(missing(na_color)) {
    na_color <- "white"
  }
  if(!is.na(str_match(string = tolower(feature), pattern = "mut"))) {
    signif_grid <- dcast(signif_df, Gene ~ Drug, value.var = "p_log10", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
    signif_grid <- signif_grid[g2p_gene_order, intersect(g2p_drug_order, colnames(signif_grid))]
    signif_grid <- signif_grid[complete.cases(signif_grid), ]
    signif_grid[signif_grid < -log10(0.05)] <- NA

    signif_grid_annot <- dcast(signif_df, Gene ~ Drug, value.var = "InG2P", fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
    signif_grid_annot <- signif_grid_annot[g2p_gene_order, intersect(g2p_drug_order, colnames(signif_grid))]
    signif_grid_annot <- signif_grid_annot[complete.cases(signif_grid_annot), ]
    signif_grid_annot[signif_grid_annot == "No"] <- ""
    signif_grid_annot[signif_grid_annot == "Yes"] <- "O"
  }
  if(!is.na(str_match(string = tolower(feature), pattern = "prot"))) {
    signif_grid <- dcast(signif_df, Gene_Protein_Change ~ Drug, value.var = "p_log10", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene_Protein_Change)
    signif_grid <- signif_grid[, intersect(g2p_drug_order, colnames(signif_grid))]

    signif_grid_annot1 <- signif_grid
    signif_grid_annot1[signif_grid_annot1 <= -log10(0.05) | is.na(signif_grid_annot1)] <- ""
    signif_grid_annot1[signif_grid_annot1 > -log10(0.05)] <- "*"

    signif_grid_annot <- dcast(signif_df, Gene_Protein_Change ~ Drug, value.var = "Direction_Curated", fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene_Protein_Change)
    signif_grid_annot <- signif_grid_annot[, intersect(g2p_drug_order, colnames(signif_grid))]
    signif_grid_annot[signif_grid_annot == "resistance"] <- "X"
    signif_grid_annot[signif_grid_annot == "sensitivity"] <- "O"
  }
  if(!is.na(str_match(string = tolower(feature), pattern = "fus"))) {
    signif_grid <- dcast(signif_df, FusionPair ~ Drug, value.var = "p_log10", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-FusionPair)
    signif_grid <- signif_grid[, intersect(g2p_drug_order, colnames(signif_grid))]
    signif_grid <- signif_grid[complete.cases(signif_grid), ]
    signif_grid[signif_grid < -log10(0.05)] <- NA

    signif_grid_annot <- dcast(signif_df, FusionPair ~ Drug, value.var = "InG2P", fill = 0) %>% `rownames<-`(.[,1]) %>% select(-FusionPair)
    signif_grid_annot <- signif_grid_annot[, intersect(g2p_drug_order, colnames(signif_grid))]
    signif_grid_annot <- signif_grid_annot[complete.cases(signif_grid_annot), ]
    signif_grid_annot[signif_grid_annot == "No"] <- ""
    signif_grid_annot[signif_grid_annot == "Yes"] <- "O"
  }
  col_fun <- colorRamp2(c(legend_breaks[1], legend_breaks[length(legend_breaks)]), c("gray97", color))
  # Legend properties list
  ht_legend_list <- list(title = legend_title, title_position = "topleft", legend_direction = "horizontal", legend_width = unit(2, "in"), border = "gray70", at = legend_breaks, labels = legend_breaks)
  ht <- Heatmap(signif_grid,
                col = col_fun,
                na_col = na_color,
                rect_gp = gpar(col = "gray70"),
                heatmap_legend_param = ht_legend_list,
                cluster_rows = FALSE, 
                cluster_columns = FALSE, 
                show_row_names = as.logical(row_names),
                row_names_side = "left",
                row_names_gp = gpar(fontsize = 10),
                column_names_side = "bottom",
                column_names_gp = gpar(fontsize = 10),
                show_column_names = FALSE,
                bottom_annotation = HeatmapAnnotation(
                  text = anno_text(colnames(signif_grid),
                                   rot = 45,
                                   gp = gpar(fontsize = 11),
                                   offset = unit(1, "npc"),
                                   just = "right"),
                  annotation_height = unit(0.85, "in")),
                  # annotation_height = max_text_width(colnames(signif_grid))),
                row_title = row_title_txt,
                row_title_gp = gpar(fontface = "bold", fontsize = 14),
                column_title = col_title, 
                column_title_gp = gpar(fontface = "bold", fontsize = 14),
                column_title_side = "bottom",
                cell_fun = function(j, i, x, y, width, height, fill) {
                  if(!is.na(str_match(string = tolower(feature), pattern = "prot"))) {
                    grid.text(signif_grid_annot1[i, j], x = x - width*0.20, y = y + height*0.07, gp = gpar(fontsize = 17))
                  }
                  if(signif_grid_annot[i, j] == "O") {
                    grid.circle(x = x - width*0.20, y = y - height*0.20, r = radius, default.units = "npc", gp = gpar(fill = "white", col = "black"))
                  }
                  if(signif_grid_annot[i, j] == "X") {
                    grid.circle(x = x - width*0.20, y = y - height*0.20, r = radius, default.units = "npc", gp = gpar(fill = "red", col = "black"))
                  }
                }
                )
  return(ht)
}
```

@@

ht_ccle2 <- MakeWilcoxonHeatmap(signif_df = ccle_signif_prot_4plot, feature = "protein changes", legend_title = "CCLE -log10(p-value)                                                              ", legend_breaks = seq(0, 15, by = 3), row_names = TRUE, color = "darkgreen", col_title = "", row_title_txt = "Gene:Protein Change", radius = 0.03, na_color = "black")
ht_ctrp2 <- MakeWilcoxonHeatmap(signif_df = ctrp_signif_prot_4plot, feature = "protein changes", legend_title = "CTRP -log10(p-value)                                                                               ", legend_breaks = seq(0, 15, by = 3), row_names = FALSE, color = "steelblue4", col_title = "Drugs", row_title_txt = "Gene:Protein Change", radius = 0.015, na_color = "black")
ht_gdsc2 <- MakeWilcoxonHeatmap(signif_df = gdsc_signif_prot_4plot, feature = "protein changes", legend_title = "GDSC -log10(p-value)", legend_breaks = seq(0, 8, by = 2), row_names = FALSE, color = "turquoise4", col_title = "", row_title_txt = "Gene:Protein Change", radius = 0.015, na_color = "black")
png(filename = "./plots/manuscript/drug_prot_heatmap.png", width = 13, height = 5, units = "in", res = 500)
draw(ht_ccle2 + ht_ctrp2 + ht_gdsc2, gap = unit(0.25, "in"), heatmap_legend_side = "top")
dev.off()
```
